### 2023-04-26 paper.tex
replaced $$ by \[

For later use: We define $C$ as the set of all cells in the assay (i.e., $C$ is the index set for the cell indices $j$).
Moreover, we define $C_i\subset C$ as the set of all those cells $j$ that have reads covering position $i$,
$$ C_i=\{j\in C: x_{ij}\neq\text{NA}\}.$$
Conversely, we define $G_j$ as the set of all the CpG positions $i$ covered by reads from cell $j$ 
$$ G_j=\{i: x_{ij}\neq\text{NA}\}.$$

@@

For each CpG position $i$, we write 
$$\overline{x}_i=\langle x_{ij} \rangle_{j\in C_i} = \frac{1}{|C_i|}\sum_{j\in C_i} x_{ij}$$ 
for the average methylation at position $i$, where $\langle\cdot\rangle$ denotes averaging, and the average runs over all the cells $j\in C_i$, i.e.
over those cells for which methylation data is available for position $i$.

@@

$$ m_{Ij} = \left< x_{ij} \right>_{i\in I\cap G_j}.$$

@@

$$ z_{Ij} = m_{Ij} - \langle m_{Ij'}\rangle_{j'\in C}.$$

As an alternative, we suggest to consider the residuals of the individual CpG methylation values $x_{ij}$ from the smoothed average $\tilde x_i$,
$$ r_{ij} = x_{ij} - \tilde x_i, $$
and averaging over these, obtaining
\begin{equation} 
r_{Ij} = \frac{\scriptsize{1}}{\scriptsize{|I\cap C_i|+1}}\sum_{i\in I\cap C_i}\left( x_{ij} - \tilde x_i \right).
\label{avgres}
\end{equation}

@@

When searching for DMRs, we compare two groups of cells, whose index sets we denote with $C_\text{A}$ and $C_\text{B}$.
For a given interval $I$, we calculate the mean each of the mean shrunken residuals $r_{ij}$ (see Eq.~(\ref{avgres})) over the cells $j$ in each of the two groups,
$$ \mu^g_I = \langle r_{Ij} \rangle_{j\in g},\qquad g=\text{A},\text{B} $$

We also calculate a variance as in Eq.(\ref{var}):
$$ v^g_I = \frac{1}{|C_\text{g}|-1}  \sum_{j\in g} \left( r_{Ij} - \mu_I^g \right)^2, 
\qquad g=\text{A},\text{B}$$

From this, we calculate Welch's t statistic as usual:
$$ t_I = \frac{\mu^\text{A}_I - \mu^\text{B}_I}{\sqrt{\frac{v^\text{A}_I}{|C_A|} + \frac{v^\text{B}_I}{|C_B|}}}.$$

@@

All the t values obtained from this permutation procedure are taken together to get an empirical null distribution.
Then, we can use this null distribution to control false discovery rate (FDR) by applying the Benjamini-Hochberg procedure in its p-value-free form:
Let us write $T$ for the set of all t values obtained from the ``real'' assignment of cells to group labels and $T_0$ from the set of all t values obtained from the shuffled assignments, i.e., the empirical null distribution.
To adjust a specific t value $t_i\in T$, we calculate
$$ p^\text{adj}_i = \frac{ \left|\left\{t_j\in T_0\mid|t_j| > |t_i| \right\}\right| \big/ |T_0|}
{ \left|\left\{t_j\in T\mid|t_j| > |t_i| \right\}\right| \big/ |T|}.$$

>>>>

For later use: We define $C$ as the set of all cells in the assay (i.e., $C$ is the index set for the cell indices $j$).
Moreover, we define $C_i\subset C$ as the set of all those cells $j$ that have reads covering position $i$,
\[ C_i=\{j\in C: x_{ij}\neq\text{NA}\}.\]
Conversely, we define $G_j$ as the set of all the CpG positions $i$ covered by reads from cell $j$ 
\[ G_j=\{i: x_{ij}\neq\text{NA}\}.\]

@@

For each CpG position $i$, we write 
\[\overline{x}_i=\langle x_{ij} \rangle_{j\in C_i} = \frac{1}{|C_i|}\sum_{j\in C_i} x_{ij}\] 
for the average methylation at position $i$, where $\langle\cdot\rangle$ denotes averaging, and the average runs over all the cells $j\in C_i$, i.e.
over those cells for which methylation data is available for position $i$.

@@

\[ m_{Ij} = \left< x_{ij} \right>_{i\in I\cap G_j}.\]

@@

\[ z_{Ij} = m_{Ij} - \langle m_{Ij'}\rangle_{j'\in C}.\]

As an alternative, we suggest to consider the residuals of the individual CpG methylation values $x_{ij}$ from the smoothed average $\tilde x_i$,
\[ r_{ij} = x_{ij} - \tilde x_i, \]
and averaging over these, obtaining
\begin{equation} 
r_{Ij} = \frac{\scriptsize{1}}{\scriptsize{|I\cap C_i|+1}}\sum_{i\in I\cap C_i}\left( x_{ij} - \tilde x_i \right).
\label{avgres}
\end{equation}

@@

When searching for DMRs, we compare two groups of cells, whose index sets we denote with $C_\text{A}$ and $C_\text{B}$.
For a given interval $I$, we calculate the mean each of the mean shrunken residuals $r_{ij}$ (see Eq.~(\ref{avgres})) over the cells $j$ in each of the two groups,
\[ \mu^g_I = \langle r_{Ij} \rangle_{j\in g},\qquad g=\text{A},\text{B}. \]

We also calculate a variance as in Eq.(\ref{var}):
\[ v^g_I = \frac{1}{|C_\text{g}|-1}  \sum_{j\in g} \left( r_{Ij} - \mu_I^g \right)^2, 
\qquad g=\text{A},\text{B}.\]

From this, we calculate Welch's t statistic as usual:
\[ t_I = \frac{\mu^\text{A}_I - \mu^\text{B}_I}{\sqrt{\frac{v^\text{A}_I}{|C_A|} + \frac{v^\text{B}_I}{|C_B|}}}.\]

@@

All the t values obtained from this permutation procedure are taken together to get an empirical null distribution.
Then, we can use this null distribution to control false discovery rate (FDR) by applying the Benjamini-Hochberg procedure in its p-value-free form:
Let us write $T$ for the set of all t values obtained from the ``real'' assignment of cells to group labels and $T_0$ from the set of all t values obtained from the shuffled assignments, i.e., the empirical null distribution.
To adjust a specific t value $t_i\in T$, we calculate
\[ p^\text{adj}_i = \frac{ \left|\left\{t_j\in T_0\mid|t_j| > |t_i| \right\}\right| \big/ |T_0|}
{ \left|\left\{t_j\in T\mid|t_j| > |t_i| \right\}\right| \big/ |T|}.\]
